Stellar Radiation is Critical for Regulating Star Formation and Driving Outflows in Low Mass Dwarf Galaxies

### 2018-07-24 ms.tex
some revisions, a new reference

Radiation from massive stars dominates the total feedback energy output of a stellar population \citep[e.g.][]{Leitherer1999,Agertz2013}, surpassing the energy ejection of supernova ($\sim 10^{51}$~erg) by two orders of magnitude. If the radiation couples effectively to the interstellar medium (ISM), it can be a substantial source of additional feedback. %Lower mass stars (size range) primarily drive photoelectric heating (cite) and H$_2$ dissociation through radiation in the far ultraviolet and Lyman-Werner bands, while more massive stars (range) emit significant amounts of ionizing radiation. 
Simulations including stellar radiation feedback followed through radiative transfer or radiation-hydrodynamics schemes have found it to be effective in regulating star formation and driving galactic winds \citep[e.g.][]{WiseAbel2012,Kim2013a,Sales2014,Oshea2015,Rosdahl2015,Ocvirk2015,Pawlik2015,Peters2017}. This occurs for four reasons: 1) heating gas and preventing the formation of cold, dense star formation regions, 2) the destruction of cold, dense gas around recently formed stars, preventing further star formation, 3) momentum input by direct absorption of UV radiation by gas and (in some cases) dust through re-emission and scattering in the infrared, and 4) lowering the typical ISM densities in which SNe occur and greatly increasing their effectiveness. % the relevance / importance of RP from dust in the IR is beyond the scope of this work...

@@

\section{Methods and Initial Conditions} \label{sec:methods}
We refer the reader to Paper I for a more detailed description of our methods, briefly summarized here. We use the adaptive mesh refinement hydrodynamics code \textsc{Enzo} \citep{Enzo2014} to evolve an idealized, isolated low mass dwarf galaxy. The galaxy is initialized as a smooth exponential disk set in hydrostatic equillibrium with a static dark matter potential \citep{Burkert1995} with $M_{\rm gas} = 2 \times 10^6~$~M$_{\odot}$, radial and vertical gas scale heights of 250~pc and 100~pc respectively, $M_{\rm vir} = 2.48 \times 10^9$~M$_{\odot}$, and a maximum physical resolution of 1.8~pc. We do not model an initial stellar population, but do include random driving from supernovae as an initial source of initial and feedback up until first star formation.

@@

\begin{figure*}
\centering
\includegraphics[width=0.49\linewidth]{sfr}
\includegraphics[width=0.49\linewidth]{mass}
\caption{The star formation rate (left) and gas and stellar mass (right) of each of our three simulations over time.}
\label{fig:sfr_mass_evolution}
\end{figure*}

We compare the resulting star formation rate (left) and gas mass properties (right) of our three simulations in Figure~\ref{fig:sfr_mass_evolution}. There is a clear, significant contrast between the runs with and without ionizing radiation. Stellar ionizing radiation leads to a factor of $\sim$ 5 reduction in the resulting SFR, as compared to the noRT run. Since the shortrad and fiducial simulations are so similar over the first $\sim$~100~Myr, it is clear that stellar radiation acts to significantly reduce the local star formation efficiency around young, massive stars. Radiation from these stars quickly ionizes and dissipates surrounding dense gas that would otherwise have gone to fuel a significant amount of additional star formation. We are unable to follow the long-term evolution of the noRT galaxy due to computational constraints. Although radiative transfer is expensive, this run is substantially more computationally expensive than the fiducial and shortrad runs due to the significantly higher star formation rate.

Looking again at the first $\sim$100~Myr of simulation time, the effects of ionizing radiation beyond our 20~pc cutoff radius are not significant. However, these two simulations begin to diverge significantly after this point. The shortrad simulation has continual, steady star formation for the entire simulation time, while the star formation rate in our fiducial run is bursty, with periods of active, low-level star formation interspersed with periods (sometimes up to 50~Myr) of no star formation. The driver of these differences is clear in the right hand panel. Our fiducial run looses a significant amount of total gas and HI mass (black and blue lines) as compared to the shortrad simulation by a factor of $\sim$2. Clearly, galactic winds and outflows are much more effective in the fiducial case, with full accounting of stellar ionizing radiation feedback, then in the shortrad simulation.

This can be confirmed by examining the metal retention fraction (the fraction of produced metals retained within the disk of the galaxy) and mass outflow rates in each simulation. As shown in Figure~\ref{fig:metal_retention}, the mass outflow rate in the fiducial run peaks at an order of magnitude higher at 0.25~R$_{\rm vir}$ than the shortrad simulation, declining only due to a comparative drop off in star formation. Given this, the fiducial simulation has an even higher mass loading factor, since the shortrad simulation has a typically higher SFR. The fiducial run is also the only simulation with any significant outflow beyond the virial radius, though we note the noRT run has not been simulated long enough for gas to have reached the virial radius. The outflow in noRT is nearly the same as the fiducial run, but it requires a 5 times higher supernova rate in this simulation to match the same outflow seen with full stellar radiation feedback. Radiation feedback, therefore allows SN to be substantially more effective in driving outflows. We provide a physical intuition for these differences in the following section.

>>>>

Radiation from massive stars dominates the total feedback energy output of a stellar population 
%mm added Abbott1982
\citep[e.g.][]{Abbott1982,Leitherer1999,Agertz2013}, surpassing the energy ejection of supernova ($\sim 10^{51}$~erg) by two orders of magnitude. If the radiation couples effectively to the interstellar medium (ISM), it can be a substantial source of additional feedback. %Lower mass stars (size range) primarily drive photoelectric heating (cite) and H$_2$ dissociation through radiation in the far ultraviolet and Lyman-Werner bands, while more massive stars (range) emit significant amounts of ionizing radiation. 
Simulations including stellar radiation feedback followed through radiative transfer or radiation-hydrodynamics schemes have found it to be effective in regulating star formation and driving galactic winds \citep[e.g.][]{WiseAbel2012,Kim2013a,Sales2014,Oshea2015,Rosdahl2015,Ocvirk2015,Pawlik2015,Peters2017}. This occurs in four ways: 1) heating 
%mm 
   of diffuse 
gas and preventing the formation of cold, dense star formation regions, 2) destruction of cold, dense gas around recently formed stars, preventing further star formation, 3) momentum input by direct absorption of UV radiation by gas and (in some cases) dust through re-emission and scattering in the infrared, and 4) lowering the typical ISM densities in which SNe occur and greatly increasing their effectiveness. % the relevance / importance of RP from dust in the IR is beyond the scope of this work...

@@

\section{Methods and Initial Conditions} \label{sec:methods}
We refer the reader to Paper I for a more detailed description of our methods, briefly summarized here. We use the adaptive mesh refinement hydrodynamics code \textsc{Enzo} \citep{Enzo2014} to evolve an idealized, isolated low mass dwarf galaxy. The galaxy is initialized as a smooth exponential disk set in hydrostatic equillibrium with a static dark matter potential \citep{Burkert1995} with $M_{\rm gas} = 2 \times 10^6~$~M$_{\odot}$, radial and vertical gas scale heights of 250~pc and 100~pc respectively, $M_{\rm vir} = 2.48 \times 10^9$~M$_{\odot}$, on a grid with a maximum physical resolution of 1.8~pc. We do not model an initial stellar population, but do include random driving from supernovae as an initial source of feedback up until the first explicit star formation in our model.

@@

\begin{figure*}
\centering
\includegraphics[width=0.49\linewidth]{sfr}
\includegraphics[width=0.49\linewidth]{mass}
\caption{The star formation rate (left) and gas and stellar mass (right) of each of our three simulations over time.
%mm
   Time bins with no star formation are left empty. 
   {\bf legend on first plot has Shortard rather than Shortrad.}}
\label{fig:sfr_mass_evolution}
\end{figure*}

We compare the resulting star formation rate (left) and gas mass properties (right) of our three simulations in Figure~\ref{fig:sfr_mass_evolution}. There is a clear, significant contrast between the runs with and without ionizing radiation. Stellar ionizing radiation leads to a factor of $\sim$ 5 reduction in the resulting SFR, as compared to the noRT run. Since the shortrad and fiducial simulations are so similar over the first $\sim$~100~Myr, it is clear that stellar radiation acts to significantly reduce the local star formation efficiency around young, massive stars. Radiation from these stars quickly ionizes and dissipates surrounding dense gas that would otherwise have gone to fuel a significant amount of additional star formation. We are unable to follow the long-term evolution of the noRT galaxy due to computational constraints 
%mm
   from the large number of stars formed. 
Although radiative transfer is computationally expensive, this run is substantially more expensive than the fiducial and shortrad runs due to the significantly higher star formation rate.

Looking again at the first $\sim$100~Myr of simulation time, the effects of ionizing radiation beyond our 20~pc cutoff radius are not significant. However, these two simulations begin to diverge significantly after this point. The shortrad simulation has continual, steady star formation for the entire simulation time, while the star formation rate in our fiducial run is bursty, with periods of active, low-level star formation interspersed with periods (sometimes up to 50~Myr) of no star formation. The driver of these differences is clear in the right hand panel. Our fiducial run loses 
%mm a significant amount of 
    a factor of $\sim 2$ more 
total gas and HI mass (black and blue lines) as compared to the shortrad simulation. 
%mm by a factor of $\sim$2. 
Clearly, galactic winds and outflows are much more effective in the fiducial case, with full accounting of stellar ionizing radiation feedback, then in the shortrad simulation.

This can be confirmed by examining the metal retention fraction (the fraction of produced metals retained within the disk of the galaxy) and mass outflow rates in each simulation. As shown in Figure~\ref{fig:metal_retention}, the mass outflow rate in the fiducial run peaks at an order of magnitude higher at 0.25~R$_{\rm vir}$ than the shortrad simulation, declining only due to a comparative drop off in star formation. Given this, the fiducial simulation has an even higher mass loading factor, since the shortrad simulation has a typically higher SFR. The fiducial run is also the only simulation with any significant outflow beyond the virial radius, though we note the noRT run has not been simulated long enough for gas to have reached the virial radius. The outflow in noRT is nearly the same as the fiducial run, but it requires a 5 times higher supernova rate in this simulation to match the same outflow seen with full stellar radiation feedback. Radiation feedback, therefore allows SN to be substantially more effective in driving outflows. We provide a physical
%mm intuition 
   explanation
for these differences in the following section.
